---
title: "BodyTemp-Blerd"
author: "BDH"
format: html
editor: visual
---

# Data Set Up

```{r}

# 1.  
library(tidyverse)
library(car)
library(gt)
library(modelsummary)
library(emmeans)
library(modelbased)
library(multcomp)
library(ggpubr)
library(cowplot)
library(lme4)
library(lmerTest)
library(outliers)
library(DHARMa)
library(statmod)
library(readr)
library(dplyr)

# 2. 
setwd("/Users/ben/Desktop/Blerd/Data Analysis")

# 3. Load your data
datum = read.csv("Blerd.csv")

# 4. Ensure categorical variables are properly set
datum$SexF = as.factor (datum$Sex)
datum$TreatmentF = as.factor(datum$Treatment)
datum$TimeF = as.factor(datum$Time)
datum$IncubatorF = as.factor (datum$Incubator)
datum$SideF = as.factor (datum$Side)
datum$PostGlucose = as.numeric (datum$PostGlucose)
datum$PostKetones = as.numeric (datum$PostKetones)
datum$PostDamage = as.numeric (datum$PostDamage)
datum$PostCORT = as.numeric (datum$PostCORT)
datum$AvgFinalTemp = as.numeric (datum$AvgFinalTemp)
datum$AvgMaxTemp = as.numeric (datum$AvgMaxTemp)

# 5. Subset data so there are no dead individuals included
datum1 = datum %>%
  filter(!Died %in% c("Died", "PIT.Death"))

# 6. Take out any of the blank cells that cause problems when graphing
datum2 = datum1 %>%
  filter(PostDamage != "", AvgMaxTemp != "")

```

# Initial Data Viz

```{r}

# 1. Correlation between the individuals that have both the final 30 min and the top 10% of body temperature values
ggplot(datum2, aes(x = AvgFinalTemp, y = AvgMaxTemp, color = TreatmentF)) +
  geom_point() +
  theme_classic()

# 2. Correlation between the average top 10% of body temperature values and post glucose
ggplot(datum2, aes(x = PostGlucose, y = AvgMaxTemp, color = TreatmentF)) +
  geom_point() +
  theme_classic()


# 3. Correlation between the average top 10% of body temperature values and post glucocorticoids
ggplot(datum2, aes(x = PostCORT, y = AvgMaxTemp, color = TreatmentF)) +
  geom_point() +
  theme_classic()

# 4. Correlation between the average top 10% of body temperature values and post DNA damage
ggplot(datum2, aes(x = PostDamage, y = AvgMaxTemp, color = TreatmentF)) +
  geom_point() +
  theme_classic()

# 5. The average top 10% of body temperature values and relationship to treatment and sex
ggplot(datum2, aes(x = SexF, y= AvgMaxTemp, color = TreatmentF)) +
  geom_point(
    position = position_jitterdodge(
      jitter.width = 0.2, 
      dodge.width = 0.5 
    )
  ) +
  theme_classic()
```

# Analysis - Correlation between maximum temperature and final 30 minutes temperature

```{r}

# Run Pearson's Correlation Test
pearson_result = cor.test(
  x = datum2$AvgFinalTemp,    # Your first continuous variable
  y = datum2$AvgMaxTemp,    # Your second continuous variable
  method = "pearson"   # Specify the Pearson method
)

# View the results
print(pearson_result)
```

**Conclusion:** Yay – they are correlated! I took this to mean I can use the ‘average maximum body temperature variable’ to correlate to other variables (Glucose, Damage, ect). This way we can include more body temperature data points.

# Analysis - Treatment and sex impact on body temperature

```{r}
#Running the models

#  1. Start with global linear mixed effects model to assess what impacts body temperature 
results = lm(AvgMaxTemp ~ TreatmentF * SexF, data = datum2, na.action = na.omit)
summary(results)
Anova(results, type = "III")

# 2. The two way interaction was not significant, so it was removed from the model
results2 = lm(AvgMaxTemp ~ TreatmentF + SexF, data = datum2, na.action = na.omit)
summary(results2)
Anova(results2, type = "III")

#Use AIC to test which one of these models has the best fit
AIC(results, results2)


#This is the final model, use 'results2' for testing the assumptions.
results_final = lm(AvgMaxTemp ~ TreatmentF + SexF, data = datum2, na.action = na.omit)
summary(results_final)
Anova(results_final, type = "III")

```

```{r}
#Assumptions of the model

# 1. Check assumptions of a linear model

# --- Extract Residuals and Fitted Values ---
#1. Extract the data frame used by the model
datum_clean <- model.frame(results_final)

# 2. Calculate and assign the fitted and residual values to this clean data frame
datum_clean$fitted <- fitted(results_final)
datum_clean$raw_resid <- residuals(results_final)
datum_clean$std_resid <- datum_clean$raw_resid / sd(datum_clean$raw_resid, na.rm = TRUE)

# 3. Use the studentized residuals method from the 'lme4' utility functions
# Although 'rstandard' doesn't work, 'residuals' is robust.
# If you really need a standardized residual, use this function:
datum_clean$std_resid <- datum_clean$raw_resid / sd(datum_clean$raw_resid, na.rm = TRUE)


# --- 2. Create the Plots ---

# A. Linearity and Homoscedasticity Plot (Residuals vs. Fitted)
ggplot(datum_clean, aes(x = fitted, y = std_resid)) +
  geom_point(alpha = 0.6) +
  # Add the ideal reference line at y=0
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  # Add a LOESS smoothing line (red line in performance)
  geom_smooth(method = "loess", se = FALSE, color = "blue", linewidth = 0.5) + 
  labs(
    title = "Linearity Check: Standardized Residuals vs. Fitted Values",
    x = "Fitted Values",
    y = "Standardized Residuals"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 10))

# B. Normality Check (Quantile-Quantile Plot)
ggplot(datum_clean, aes(sample = std_resid)) +
  # Create the Q-Q plot
  stat_qq() + 
  # Add the reference line (the ideal normal distribution line)
  stat_qq_line(color = "red") + 
  labs(
    title = "Normality Check: Q-Q Plot of Standardized Residuals",
    x = "Theoretical Quantiles",
    y = "Standardized Residuals"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 10))

# C. Scale-Location Plot (Homogeneity of Variance Check)
# This uses the square root of the absolute standardized residuals
ggplot(datum_clean, aes(x = fitted, y = sqrt(abs(std_resid)))) +
  geom_point(alpha = 0.6) +
  # Add the ideal flat reference line
  geom_smooth(method = "loess", se = FALSE, color = "blue", linewidth = 0.5) + 
  labs(
    title = "Homogeneity of Variance Check: Scale-Location Plot",
    x = "Fitted Values",
    y = expression(sqrt(abs("Standardized Residuals")))
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 10))

#D. Histogram of residuals
# Calculate mean and standard deviation of residuals for the density curve
resid_mean <- mean(datum_clean$std_resid, na.rm = TRUE)
resid_sd <- sd(datum_clean$std_resid, na.rm = TRUE)

ggplot(datum_clean, aes(x = std_resid)) +
  # 1. Create the Histogram (using density scale)
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 15, # Adjust bin count for best visualization
                 color = "black", 
                 fill = "lightblue") +
  
  # 2. Overlay the Density of the Residuals (Green Line)
  geom_density(color = "darkgreen", linewidth = 1) +
  
  # 3. Overlay the Theoretical Normal Density Curve (Red Dashed Line)
  stat_function(fun = dnorm, 
                args = list(mean = resid_mean, sd = resid_sd), 
                color = "red", 
                linewidth = 1,
                linetype = "dashed") +
  
  labs(
    title = "Residual Normality Check: Histogram with Density Curve",
    x = "Standardized Residuals (Z-Scores)",
    y = "Density"
  ) +
  theme_minimal()
```

**Conclusions:** Based on the residuals from the model, I think the assumptions are met and we can continue with analysis using this model.

**Next steps**: No further steps

# Manuscript Graph

```{r}
# 1. Add the means and 95% CIs to the dataframe for the variables we are interested in 

datum2_with_summary = datum2 %>%
  
  # 1. Group the data by the interaction terms
  group_by(TreatmentF) %>%
  
  # 2. Calculate the group-level statistics (Mean, CI, etc.)
  mutate(
    N_group = n(),   
    Mean_Temp = mean(AvgMaxTemp, na.rm = TRUE),
    SD_Temp = sd(AvgMaxTemp, na.rm = TRUE),
    SE_Temp = SD_Temp / sqrt(N_group),
    CI_Temp = 1.96 * SE_Temp,
    CI_Upper_Group = Mean_Temp + CI_Temp,
    CI_Lower_Group = Mean_Temp - CI_Temp
  ) %>%
  ungroup()
  #View the dataframe that was created 
head(datum2_with_summary)


# 2. Cleaning up data frame before plotting
  # Define the consistent dodge width and rename the data frame for plotting
plot_data = datum2_with_summary 

  #Order the time points so that Pre and Post are graphed in the correct order on the plot
plot_data$TreatmentF = as.factor(plot_data$TreatmentF)

# 3. Plotting the graph

plot = ggplot(data = plot_data, aes(x = TreatmentF, y = AvgMaxTemp, color = TreatmentF)) + 
  geom_jitter(
    size = 3.5,
    alpha = 0.5,
    position = position_jitter(width = 0.2) 
  ) +
  
  geom_point(
    aes(y = Mean_Temp, group = TreatmentF),
    size = 7.5, 
    stroke = 1,
    color = "black" 
  ) +
  
  geom_errorbar(
    aes(ymin = CI_Lower_Group, ymax = CI_Upper_Group, group = TreatmentF),
    width = 0.2,
    linewidth = 1,
    color = "black",
  ) +

  geom_line(
    aes(y = Mean_Temp, group = TreatmentF),
    linewidth = 1.5,
  ) +
  
  scale_color_manual(
    values = c(
      "Control" = "#AAB8C2",         
      "Heat" = "firebrick4"     
    )
  ) +
  
  labs(
    x = "Treatment",
    y = "Body Temperature (°C)",
    color = "Treatment Group"
  ) +  
  theme_classic() +
  
theme(
  axis.title = element_text(face = "bold", size = 30),
  axis.text = element_text(size = 25),
  strip.text = element_text(size = 20),
  panel.background = element_rect(fill = "transparent", color = NA), 
  plot.background = element_rect(fill = "transparent", color = NA),
  axis.line = element_line(linewidth = 1.5, colour = "black"),
  legend.position = "none"
)


# 4. Adding signifigance and sample size

plot1 =  plot +
  annotate(
    "text", 
    x = c(1, 2), 
    y = c(46.6, 47.8), 
    label = c("a", "b"),  
    size = 12,
    color = "black") +
   
   # Add a new layer for the sample sizes
  annotate(
    "text",
    x = c(1, 2), 
    y = c(45.6, 46.8), 
    # Add your sample sizes here
    label = c("(26)", "(27)"), 
    size = 10, # Make the font size slightly smaller
    color = "black")

# 5. Save the graph
ggsave(
  filename = "TempPlot.png", 
  plot = plot1,                       
  width = 6.49,                              
  height = 6.49,  
  units = "in",
  dpi = 500,
  bg = "transparent"
)
```

# Manuscript Graph - With sex

```{r}
#Manuscript Graph with Sex

# 1. Add the means and 95% CIs to the dataframe for the variables we are interested in 

datum2_with_summary = datum2 %>%
  
  # 1. Group the data by the interaction terms
  group_by(TreatmentF) %>%
  
  # 2. Calculate the group-level statistics (Mean, CI, etc.)
  mutate(
    N_group = n(),   
    Mean_Temp = mean(AvgMaxTemp, na.rm = TRUE),
    SD_Temp = sd(AvgMaxTemp, na.rm = TRUE),
    SE_Temp = SD_Temp / sqrt(N_group),
    CI_Temp = 1.96 * SE_Temp,
    CI_Upper_Group = Mean_Temp + CI_Temp,
    CI_Lower_Group = Mean_Temp - CI_Temp
  ) %>%
  ungroup()
  #View the dataframe that was created 
head(datum2_with_summary)


# 2. Cleaning up data frame before plotting
  # Define the consistent dodge width and rename the data frame for plotting
plot_data = datum2_with_summary 

  #Order the time points so that Pre and Post are graphed in the correct order on the plot
plot_data$TreatmentF = as.factor(plot_data$TreatmentF)

# 3. Plotting the graph

plot = ggplot(data = plot_data, aes(x = TreatmentF, y = AvgMaxTemp, color = TreatmentF)) + 
  geom_jitter(
    aes(shape = SexF),
    size = 3.5,
    alpha = 0.5,
    position = position_jitter(width = 0.2) 
  ) +
  
  geom_point(
    aes(y = Mean_Temp, group = TreatmentF),
    shape = 18,
    size = 7.5, 
    stroke = 1,
    color = "black" 
  ) +
  
  geom_errorbar(
    aes(ymin = CI_Lower_Group, ymax = CI_Upper_Group, group = TreatmentF),
    width = 0.2,
    linewidth = 1,
    color = "black",
  ) +

  geom_line(
    aes(y = Mean_Temp, group = TreatmentF),
    linewidth = 1.5,
  ) +
  
  scale_color_manual(
    values = c(
      "Control" = "#AAB8C2",         
      "Heat" = "firebrick4"     
    )
  ) +
  
  labs(
    x = "Treatment",
    y = "Body Temperature (°C)",
    color = "Treatment Group"
  ) +  
  theme_classic() +
  
theme(
  axis.title = element_text(face = "bold", size = 30),
  axis.text = element_text(size = 25),
  strip.text = element_text(size = 20),
  panel.background = element_rect(fill = "transparent", color = NA), 
  plot.background = element_rect(fill = "transparent", color = NA),
  axis.line = element_line(linewidth = 1.5, colour = "black"),
  legend.position = "none"
)


# 4. Adding signifigance and sample size

plot1 =  plot +
  annotate(
    "text", 
    x = c(1, 2), 
    y = c(46.6, 47.8), 
    label = c("a", "b"),  
    size = 12,
    color = "black") +
   
   # Add a new layer for the sample sizes
  annotate(
    "text",
    x = c(1, 2), 
    y = c(45.6, 46.8), 
    # Add your sample sizes here
    label = c("(26)", "(27)"), 
    size = 10, # Make the font size slightly smaller
    color = "black")

# 5. Save the graph
ggsave(
  filename = "TempPlotSex.png", 
  plot = plot1,                       
  width = 6.49,                              
  height = 6.49,  
  units = "in",
  dpi = 500,
  bg = "transparent"
)
```
