---
title: "DNA Damage - Blerd"
author: "BDH"
format: html
editor: visual
---

## Data Set Up

```{r}

# 1.  
library(tidyverse)
library(car)
library(gt)
library(modelsummary)
library(emmeans)
library(modelbased)
library(multcomp)
library(ggpubr)
library(cowplot)
library(lme4)
library(lmerTest)
library(outliers)
library(DHARMa)
library(statmod)
library(readr)
library(dplyr)
library(EnvStats)
library(performance)
library(see)
library(patchwork)

# 2. 
setwd("/Users/ben/Desktop/Blerd/Data Analysis")

# 3. Load your data
datum = read.csv("Blerd.csv")
datumlong = read.csv("BlerdData_long.csv")

# 4. Ensure categorical variables are properly set
datum$SexF = as.factor (datum$Sex)
datum$TreatmentF = as.factor(datum$Treatment)
datum$TimeF = as.factor(datum$Time)
datum$IncubatorF = as.factor (datum$Incubator)
datum$SideF = as.factor (datum$Side)
datum$PostDamage = as.numeric (datum$PostDamage)
datum$PreDamage = as.numeric (datum$PreDamage)

datumlong$SexF = as.factor (datumlong$Sex)
datumlong$TreatmentF = as.factor(datumlong$Treatment)
datumlong$TimepointF = as.factor(datumlong$Timepoint)
datumlong$Damage = as.numeric (datumlong$Damage)

# 5. Subset data so there are no dead individuals included
datum1 = datum %>%
  filter(!Died %in% c("Died", "PIT.Death"))

datumlong1 = datumlong %>%
  filter(!Died %in% c("Died", "PIT.Death"))
datumlong1 = datumlong1[-47, ]

# 6. Take out any of the blank cells that cause problems when graphing
datum2 = datum1 %>%
  filter(PostDamage != "", PostDamage != "Dead", AvgMaxTemp != "", AvgMaxTemp != "Dead")
datum2$TreatmentF = as.factor(datum2$Treatment)

datumlong2 = datumlong1 %>%
  filter(Damage != "", Damage != "Dead")
```

# Initial Data Viz

```{r}
# 1. Damage across timepoints and relationship to treatment and sex
ggplot(datumlong2, aes(x = TimepointF, y = Damage, color = SexF, shape = TreatmentF)) +
  geom_point(position = position_jitterdodge(
      jitter.width = 0.2, 
      dodge.width = 0.5 )) +
  theme_classic() 

# 2. Correlation between the average top 10% of body temperature values and post damage
ggplot(datum2, aes(x = AvgMaxTemp, y = PostDamage, color = TreatmentF)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  theme_classic()

```

# Summary Stats

```{r}
# 1. Averages for all variables of interest
table = datumlong2 %>%
  group_by(TreatmentF, TimepointF, SexF) %>%
  summarise(
    meanDamage = mean(Damage, na.rm = TRUE),
    sdDamage = sd(Damage, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

print(table)
```

# Analysis - Damage across time points

```{r}

# 1. Running the models

  #Start with global linear mixed effects model to assess damage levels
results = lmer(Damage ~ TreatmentF * SexF * TimepointF + (1|ID),
               data = datumlong2, na.action = na.omit, REML = FALSE)
summary(results)
Anova(results, type = "III")

# 2. The three way interaction was not significant, so it was removed and all two way interactions are included
results2 = lmer(Damage ~ TreatmentF * SexF + TreatmentF * TimepointF +
                  SexF * TimepointF + (1|ID), data = datumlong2, na.action = na.omit,              REML = FALSE)
summary(results2)
Anova(results2, type = "III")

# 3. The Sex*Timepoint interaction was not significant, so it was removed from the model
results3 = lmer(Damage ~ TreatmentF * SexF + TreatmentF * TimepointF +
                  (1|ID), data = datumlong2, na.action = na.omit, REML = FALSE)
summary(results3)
Anova(results3, type = "III")

# 4. The Treatment*Sex interaction was not significant, so it was removed from the model
results4 = lmer(Damage ~ TreatmentF * TimepointF + SexF + (1|ID), data = datumlong2,
                na.action = na.omit, REML = FALSE)
summary(results4)
Anova(results4, type = "III")

# 5. The Treatment*Sex interaction was p=0.052 , so to ensure that is not significant, we will remove the interaction from the model and see if the AIC score improves
results5 = lmer(Damage ~ TreatmentF + TimepointF + SexF + (1|ID), data = datumlong2,
                na.action = na.omit, REML = FALSE)
summary(results5)
Anova(results5, type = "III")

#Use AIC to test which one of these models has the best fit
AIC(results, results2, results3, results4, results5)


#According to AIC scores, use 'results4' for testing the assumptions

results_final = lmer(Damage ~ TreatmentF * TimepointF + SexF + (1|ID), data = datumlong2,
                na.action = na.omit, REML = TRUE)
summary(results_final)
Anova(results_final, type = "III")
```

```{r}
#Assumptions of the model


# 1. Check the VIF of this model
vif(results_final)
  #VIF = 1: No multicollinearity
  #VIF between 1 and 5: Generally considered acceptable or moderate multicollinearity
  #VIF > 5 (or > 10): Indicates potentially high multicollinearity

# 2. Check other assumptions of a linear model

# --- Extract Residuals and Fitted Values ---
#1. Extract the data frame used by the model
datum_clean = model.frame(results_final)

# 2. Calculate and assign the fitted and residual values to this clean data frame
datum_clean$fitted <- fitted(results_final)
datum_clean$raw_resid <- residuals(results_final)
datum_clean$std_resid <- datum_clean$raw_resid / sd(datum_clean$raw_resid, na.rm = TRUE)

# 3. Use the studentized residuals method from the 'lme4' utility functions
# Although 'rstandard' doesn't work, 'residuals' is robust.
# If you really need a standardized residual, use this function:
datum_clean$std_resid <- datum_clean$raw_resid / sd(datum_clean$raw_resid, na.rm = TRUE)


# --- 2. Create the Plots ---

# A. Linearity and Homoscedasticity Plot (Residuals vs. Fitted)
ggplot(datum_clean, aes(x = fitted, y = std_resid)) +
  geom_point(alpha = 0.6) +
  # Add the ideal reference line at y=0
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  # Add a LOESS smoothing line (red line in performance)
  geom_smooth(method = "loess", se = FALSE, color = "blue", linewidth = 0.5) + 
  labs(
    title = "Linearity Check: Standardized Residuals vs. Fitted Values",
    x = "Fitted Values",
    y = "Standardized Residuals"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 10))

# B. Normality Check (Quantile-Quantile Plot)
ggplot(datum_clean, aes(sample = std_resid)) +
  # Create the Q-Q plot
  stat_qq() + 
  # Add the reference line (the ideal normal distribution line)
  stat_qq_line(color = "red") + 
  labs(
    title = "Normality Check: Q-Q Plot of Standardized Residuals",
    x = "Theoretical Quantiles",
    y = "Standardized Residuals"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 10))

# C. Scale-Location Plot (Homogeneity of Variance Check)
# This uses the square root of the absolute standardized residuals
ggplot(datum_clean, aes(x = fitted, y = sqrt(abs(std_resid)))) +
  geom_point(alpha = 0.6) +
  # Add the ideal flat reference line
  geom_smooth(method = "loess", se = FALSE, color = "blue", linewidth = 0.5) + 
  labs(
    title = "Homogeneity of Variance Check: Scale-Location Plot",
    x = "Fitted Values",
    y = expression(sqrt(abs("Standardized Residuals")))
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 10))

#D. Histogram of residuals
# Calculate mean and standard deviation of residuals for the density curve
resid_mean <- mean(datum_clean$std_resid, na.rm = TRUE)
resid_sd <- sd(datum_clean$std_resid, na.rm = TRUE)

ggplot(datum_clean, aes(x = std_resid)) +
  # 1. Create the Histogram (using density scale)
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 15, # Adjust bin count for best visualization
                 color = "black", 
                 fill = "lightblue") +
  
  # 2. Overlay the Density of the Residuals (Green Line)
  geom_density(color = "darkgreen", linewidth = 1) +
  
  # 3. Overlay the Theoretical Normal Density Curve (Red Dashed Line)
  stat_function(fun = dnorm, 
                args = list(mean = resid_mean, sd = resid_sd), 
                color = "red", 
                linewidth = 1,
                linetype = "dashed") +
  
  labs(
    title = "Residual Normality Check: Histogram with Density Curve",
    x = "Standardized Residuals (Z-Scores)",
    y = "Density"
  ) +
  theme_minimal()
```

**Conclusion:** Histogram looks good. Overall, this model is a good fit.

**Next Steps:** Continue to post-hoc analysis.

```{r}
emms_interaction = emmeans(results_final, ~ TreatmentF * TimepointF)
interaction_pairs = pairs(emms_interaction, adjust = "sidak")
print(interaction_pairs)

cld_output = cld(emms_interaction,
                  adjust = "sidak",   
                  Letters = letters)

print(cld_output)

```

# Analysis - Damage across time % change

```{r}

# 1. Making the column

datum2$DeltaDamage = datum2$PostDamage/datum2$PreDamage

# 2. Running the models

  #Start with global linear model to assess damage levels
results = lm(DeltaDamage ~ TreatmentF * SexF, data = datum2, na.action = na.omit)
summary(results)
Anova(results, type = "III")

  #The interaction was not significant, so it was removed from the model
results2 = lm(DeltaDamage ~ TreatmentF + SexF, data = datum2, na.action = na.omit)
summary(results2)
Anova(results2, type = "III")

  #Neither Sex nor Treatment affected the outcome, so each main effect was ran by itself
results3 = lm(DeltaDamage ~ SexF, data = datum2, na.action = na.omit)
summary(results3)
Anova(results3, type = "III")

  #Neither Sex nor Treatment affected the outcome, so each main effect was ran by itself
results4 = lm(DeltaDamage ~ TreatmentF, data = datum2, na.action = na.omit)
summary(results4)
Anova(results4, type = "III")

```

The outcomes here are not any different than when we include time point, and so I would like to keep time point in the model.

# Analysis - Body Temperature and Damage Correlation

```{r}

# 1. Run Pearson's Correlation Test for both heated and control groups
pearson_result = cor.test(
  x = datum2$PostDamage,    
  y = datum2$AvgMaxTemp,    
  method = "pearson"  
)
  # View the results
print(pearson_result)


# 2.  Run Pearson's Correlation Test heated group
datumHeated = datum2 %>% 
 filter(TreatmentF == "Heat")

pearson_resultHeat = cor.test(
  x = datumHeated$PostDamage,   
  y = datumHeated$AvgMaxTemp,   
  method = "pearson"  
)
  # View the results
print(pearson_resultHeat)


# 2.  Run Pearson's Correlation Test control group
datumControl = datum2 %>% 
 filter(TreatmentF == "Control")

pearson_resultControl = cor.test(
  x = datumControl$PostDamage,    
  y = datumControl$AvgMaxTemp,    
  method = "pearson"  
)
  # View the results
print(pearson_resultControl)
```

Nothing was significant here, which makes sense based on previous models

# Manuscript Graph

```{r}
# 1. Add the means and 95% CIs to the dataframe for the variables we are interested in 

datumlong2_with_summary = datumlong2 %>%
  
  # 1. Group the data by the interaction terms
  group_by(TreatmentF, TimepointF) %>%
  
  # 2. Calculate the group-level statistics (Mean, CI, etc.)
  mutate(
    N_group = n(),   
    Mean_Damage_Group = mean(Damage, na.rm = TRUE),
    SD_Damage_Group = sd(Damage, na.rm = TRUE),
    SE_Damage_Group = SD_Damage_Group / sqrt(N_group),
    CI_Margin_Damage = 1.96 * SE_Damage_Group,
    CI_Upper_Group = Mean_Damage_Group + CI_Margin_Damage,
    CI_Lower_Group = Mean_Damage_Group - CI_Margin_Damage
  ) %>%
  ungroup()
  #View the dataframe that was created 
head(datumlong2_with_summary)


# 2. Cleaning up data frame before plotting
  # Define the consistent dodge width and rename the data frame for plotting
DODGE_WIDTH = 0.4 
plot_data = datumlong2_with_summary 

  #Order the time points so that Pre and Post are graphed in the correct order on the plot
plot_data$TreatmentF = as.factor(plot_data$TreatmentF)
plot_data$TimepointF = factor(plot_data$TimepointF, levels = c("Pre", "Post")) 

# 3. Plotting the graph

plot = ggplot(data = plot_data, aes(x = TimepointF, y = Damage, color = TreatmentF)) + 
  geom_jitter(
    size = 3.5,
    alpha = 0.5,
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = DODGE_WIDTH) 
  ) +
  
  geom_point(
    aes(y = Mean_Damage_Group, group = TreatmentF),
    size = 7.5, 
    stroke = 1,
    color = "black", 
    position = position_dodge(width = DODGE_WIDTH) 
  ) +
  
  geom_errorbar(
    aes(ymin = CI_Lower_Group, ymax = CI_Upper_Group, group = TreatmentF),
    width = 0.2,
    linewidth = 1,
    color = "black",
    position = position_dodge(width = DODGE_WIDTH) 
  ) +

  geom_line(
    aes(y = Mean_Damage_Group, group = TreatmentF),
    linewidth = 1.5,
    position = position_dodge(width = DODGE_WIDTH) 
  ) +
  
  scale_color_manual(
    values = c(
      "Control" = "#AAB8C2",         
      "Heat" = "firebrick4"     
    )
  ) +
  
  labs(
    x = "Time Point",
    y = "DNA Damage (AU)",
    color = "Treatment Group"
  ) +  
  theme_classic() +
  
  theme(
    axis.title = element_text(face = "bold", size = 30),
    axis.text = element_text(size = 25),
    legend.justification = c(0.5, 0.5),
    legend.title = element_text(size = 28),
    legend.text = element_text(size = 25),
    strip.text = element_text(size = 20),
    panel.background = element_rect(fill = "transparent", color = NA), 
    plot.background = element_rect(fill = "transparent", color = NA),  
    legend.background = element_rect(fill = "transparent", color = NA), 
    legend.box.background = element_rect(fill = "transparent", color = NA),
    axis.line = element_line(linewidth = 1.5, colour = "black")
  ) 

# 4. Adding signifigance and sample size

plot1 = plot +
  annotate(
    "text", 
    x = c(.9, 1.1, 1.9, 2.1), 
    y = c(2.1 + .3, 2.5 + .3, 3.1 + .3, 2 + .3), 
    label = c("a", "a", "a", "a"),  
    size = 10,
    color = "black") +
   
   # Add a new layer for the sample sizes
  annotate(
    "text",
    x = c(.9, 1.1, 1.9, 2.1),
    # Adjust the y-coordinates to be slightly lower
    y = c(2.1, 2.5, 3.1, 2), 
    # Add your sample sizes here
    label = c("(34)", "(42)", "(36)", "(35)"), 
    size = 7.5, # Make the font size slightly smaller
    color = "black")

# 5. Save the graph
ggsave(
  filename = "DamagePlot.png", 
  plot = plot1,                       
  width = 11.04,                              
  height = 6.49,  
  units = "in",
  dpi = 300,
  bg = "transparent"
)
```

# Manuscript Graph - With Sex

```{r}
# 1. Add the means and 95% CIs to the dataframe for the variables we are interested in 

datumlong2_with_summary = datumlong2 %>%
  
  # 1. Group the data by the interaction terms
  group_by(TreatmentF, TimepointF) %>%
  
  # 2. Calculate the group-level statistics (Mean, CI, etc.)
  mutate(
    N_group = n(),   
    Mean_Damage_Group = mean(Damage, na.rm = TRUE),
    SD_Damage_Group = sd(Damage, na.rm = TRUE),
    SE_Damage_Group = SD_Damage_Group / sqrt(N_group),
    CI_Margin_Damage = 1.96 * SE_Damage_Group,
    CI_Upper_Group = Mean_Damage_Group + CI_Margin_Damage,
    CI_Lower_Group = Mean_Damage_Group - CI_Margin_Damage
  ) %>%
  ungroup()
  #View the dataframe that was created 
head(datumlong2_with_summary)


# 2. Cleaning up data frame before plotting
  # Define the consistent dodge width and rename the data frame for plotting
DODGE_WIDTH = 0.4 
plot_data = datumlong2_with_summary 

  #Order the time points so that Pre and Post are graphed in the correct order on the plot
plot_data$TreatmentF = as.factor(plot_data$TreatmentF)
plot_data$TimepointF = factor(plot_data$TimepointF, levels = c("Pre", "Post")) 

# 3. Plotting the graph

plot = ggplot(data = plot_data, aes(x = TimepointF, y = Damage, color = TreatmentF)) + 

  geom_jitter(
    aes(shape = SexF, group = TreatmentF), 
    size = 3.5,
    alpha = 0.5,
    position = position_jitterdodge(
      jitter.width = 0.2, 
      dodge.width = DODGE_WIDTH
  ) 
) +
  
  geom_point(
    aes(y = Mean_Damage_Group, group = TreatmentF),
    size = 7.5, 
    shape = 18,
    stroke = 1,
    color = "black", 
    position = position_dodge(width = DODGE_WIDTH) 
  ) +
  
  geom_errorbar(
    aes(ymin = CI_Lower_Group, ymax = CI_Upper_Group, group = TreatmentF),
    width = 0.2,
    linewidth = 1,
    color = "black",
    position = position_dodge(width = DODGE_WIDTH) 
  ) +

  geom_line(
    aes(y = Mean_Damage_Group, group = TreatmentF),
    linewidth = 1.5,
    position = position_dodge(width = DODGE_WIDTH) 
  ) +
  
  scale_color_manual(
    values = c(
      "Control" = "#AAB8C2",         
      "Heat" = "firebrick4"     
    )
  ) +
  
  labs(
    x = "Time Point",
    y = "DNA Damage (AU)",
    color = "Treatment Group"
  ) +  
 
  guides(shape = "none") + 
  
  theme_classic() +
  
  
  theme(
    axis.title = element_text(face = "bold", size = 30),
    axis.text = element_text(size = 25),
    legend.justification = c(0.5, 0.5),
    legend.title = element_text(size = 28),
    legend.text = element_text(size = 25),
    strip.text = element_text(size = 20),
    panel.background = element_rect(fill = "transparent", color = NA), 
    plot.background = element_rect(fill = "transparent", color = NA),  
    legend.background = element_rect(fill = "transparent", color = NA), 
    legend.box.background = element_rect(fill = "transparent", color = NA),
    axis.line = element_line(linewidth = 1.5, colour = "black")
  ) 

# 4. Adding signifigance and sample size

plot1 = plot +
  annotate(
    "text", 
    x = c(.9, 1.1, 1.9, 2.1), 
    y = c(2.1 + .3, 2.5 + .3, 3.1 + .3, 2 + .3), 
    label = c("a", "a", "a", "a"),  
    size = 10,
    color = "black") +
   
   # Add a new layer for the sample sizes
  annotate(
    "text",
    x = c(.9, 1.1, 1.9, 2.1),
    # Adjust the y-coordinates to be slightly lower
    y = c(2.1, 2.5, 3.1, 2), 
    # Add your sample sizes here
    label = c("(34)", "(42)", "(36)", "(35)"), 
    size = 7.5, # Make the font size slightly smaller
    color = "black")

# 5. Save the graph
ggsave(
  filename = "DamagePlotSex.png", 
  plot = plot1,                       
  width = 11.04,                              
  height = 6.49,  
  units = "in",
  dpi = 300,
  bg = "transparent"
)
```
